---
title: "Costa Rica Carbon Stats Output"
author: "Kenna Rewcastle"
date: "4/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(dplyr)
library(MASS)
library(car)
library(agricolae)
library(tidyr)
library(reshape2)
library(nlme)
library(RLRsim)
```


# Delta 13C-CO2 Models

```{r read_13C_CO2_data, include=FALSE}
  ## Read in data
  dat<-read.csv("Costa Rica Master sheet annotated.csv")
  dat<-filter(dat,dat$Exclusion!="NA")
  dat$Exclusion<-as.factor(dat$Exclusion)
  dat$Tree_ID<-substr(dat$core_ID,1,5)

  ## Separate starch and leaf data
  starch_CO2<-filter(dat,dat$Isotope_label=="S") 
  leaf_CO2<-filter(dat,dat$Isotope_label=="L")
```

## **Leaf Substrate**: Step-wise AIC Model Selection for 13C-CO2
```{r leaf_13C_CO2_variables, include=FALSE}
    leaf_d13<-leaf_CO2$d13_d9
    Exclusion<-leaf_CO2$Exclusion
    Tree_Species<-leaf_CO2$Tree_species
```

```{r leaf_13C_CO2_stepAIC}
leaf_CO2_AIC<-stepAIC(object=aov(lm(leaf_d13~Exclusion+Tree_Species+Exclusion*Tree_Species)),direction="both",na.rm=TRUE)
```

```{r leaf_13C_CO2_AIC_anova_table}
leaf_CO2_AIC$anova
```

Model selection indicates that **rhizosphere exclusion, tree species, and exclusion x tree species** should all be included in model explaining **13C-CO2 efflux** from cores labeled with **leaf substrate**.

### Type II ANOVA Results
```{r leaf_13C_CO2_ANOVA}
leafMod<-lm(d13_d9~Exclusion*Tree_species,data=leaf_CO2,na.action=na.omit)
leaf_CO2_fixed<-Anova(leafMod,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r leaf_nested_13CO2}
leaf_CO2_mixed_mod<-lme(d13_d9~Exclusion*Tree_species,random=~1|Tree_ID,data=leaf_CO2,na.action=na.omit,method="ML")
leaf_CO2_fixed_mod<-lm(d13_d9~Exclusion*Tree_species,data=leaf_CO2,na.action=na.omit)
anova(leaf_CO2_mixed_mod,leaf_CO2_fixed_mod)
```

This output indicates that the AIC score for the mixed model is ~1.5 points lower than the fixed effects model, while BIC is virtually identical, and the p-value is 0.0639.TL;DR: Unclear whether tree ID random intercept is necessary. Evaluate other models first.

```{r leaf_CO2_mixed_mod}
summary(leaf_CO2_mixed_mod)
```

```{r leaf_CO2_fixed_mod}
summary(leaf_CO2_fixed_mod)
```

### Tukey's HSD Results
```{r leaf_13C_CO2_TukeyHSD}
 tx<-with(leaf_CO2, interaction(Exclusion, Tree_species))
  amod<-aov(d13_d9 ~ tx, data=leaf_CO2)
  leaf_CO2_Tukey<-HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

## **Starch Substrate**: Step-wise AIC Model Selection for 13C-CO2
```{r starch_13C_CO2_variables, include=FALSE}
    starch_d13<-starch_CO2$d13_d2
    Exclusion<-starch_CO2$Exclusion
    Tree_Species<-starch_CO2$Tree_species
```

```{r starch_13C_CO2_stepAIC}
starch_CO2_AIC<-stepAIC(object=aov(lm(starch_d13~Exclusion+Tree_Species+Exclusion*Tree_Species)),direction="both",na.rm=TRUE)
```

```{r starch_13C_CO2_AIC_anova_table}
starch_CO2_AIC$anova
```

```{r starch_13C_CO2_AIC_extraction}
print(extractAIC(aov(lm(starch_d13~Exclusion))))
print(extractAIC(aov(lm(starch_d13~Exclusion+Tree_Species))))
print(extractAIC(aov(lm(starch_d13~Exclusion*Tree_Species))))
```

Model selection indicates that **rhizosphere exclusion and tree species** should be included in model explaining **13C-CO2 efflux** from cores labeled with **starch substrate**. Model with only rhizosphere exclusion and model with rhizosphere exclusion and tree species are separated by only 0.7 AIC points, so both models are included in "best family" of models. Will choose to include tree species as a parameter to match structure of leaf models.

### Type II ANOVA Results
```{r starch_13C_CO2_ANOVA}
starchMod<-lm(d13_d2~Exclusion+Tree_species,data=starch_CO2)
  Anova(starchMod,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r starch_nested_13CO2}
starch_CO2_mixed_mod<-lme(d13_d2~Exclusion+Tree_species,random=~1|Tree_ID,data=starch_CO2,na.action=na.omit,method="ML")
starch_CO2_fixed_mod<-lm(d13_d2~Exclusion+Tree_species,data=starch_CO2,na.action=na.omit)
anova(starch_CO2_mixed_mod,starch_CO2_fixed_mod)
```

This output indicates that the AIC score for the mixed model is ~2.6 points lower than the fixed effects model, while BIC scores are only 1 point apart, and the p-value is 0.0305.TL;DR: Slightly more decisive that the mixed effects model is best.

```{r starch_CO2_mixed_mod}
summary(starch_CO2_mixed_mod)
```

```{r starch_CO2_fixed_mod}
summary(starch_CO2_fixed_mod)
```


### Tukey's HSD Results
```{r starch_13C_CO2_TukeyHSD}
  tx<-with(starch_CO2, interaction(Exclusion, Tree_species))
  amod<-aov(d13_d2 ~ tx, data=starch_CO2)
  HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

# Delta 13C-PLFA Models
```{r read_13C_PLFA_data, include=FALSE}
  ## Read in data
  data<-read.csv("FINAL_PLFA_with_metadata.csv")
  data<-data[,2:15] # Gets rid of ID column
  data<-filter(data,Microbial_Group!="standard")
  data$Rhizosphere_Manipulation<-as.factor(data$Rhizosphere_Manipulation)
  data$umol_FAME<-data$nmol_FAME_per_g_soil/1000
  
  ## Simplify the microbial groupings
  data_trim<-data.frame("Sample_Name"=data[,1],"Tree"=data[,2],"Rhizosphere_Manipulation"=data[,3],"Substrate_Type"=data[,4],"Microbial_Group"=data[,14],"d13"=data[,9])
  
  data_wide<-dcast(data_trim,Sample_Name+Tree+Rhizosphere_Manipulation+Substrate_Type ~ Microbial_Group, value.var="d13", fun.aggregate=mean, na.rm=TRUE)
  
  is.nan.data.frame <- function(x)
    do.call(cbind, lapply(x, is.nan))
  data_wide[is.nan(data_wide)] <- NA
  
      ### Take mean of narrow groups to yield FUNGI, BACTERIA, ALL_MICROBE groups
      for (i in 1:nrow(data_wide)) {
        data_wide$BACTERIA[i]<-mean(c(data_wide$bacteria[i],data_wide$`gram+bacteria`[i],data_wide$`gram-bacteria`[i]),na.rm=TRUE)
        data_wide$FUNGI[i]<-mean(c(data_wide$AMF[i],data_wide$fungi[i]),na.rm=TRUE)
        data_wide$ALL_MICROBES[i]<-mean(c(data_wide$AMF[i],data_wide$bacteria[i],data_wide$fungi[i],data_wide$general[i],data_wide$`gram-bacteria`[i],data_wide$`gram+bacteria`[i],data_wide$protozoa[i]),na.rm=TRUE)
      }

  data_wide<-data_wide[,c(1,2,3,4,5,12,13,14)] # Trim down to broader groups
  
  data<-melt(data_wide,value.name="d13",id.vars=c("Sample_Name","Rhizosphere_Manipulation","Tree","Substrate_Type"))
  names(data)[5]<-"Microbial_Group"
  
  ## Add a column for tree ID
  data$Tree_ID<-substr(data$Sample_Name,1,5)
  
  ## Separate leaf and starch datasets
  leaf_PLFA<-filter(data,Substrate_Type=="L")
  starch_PLFA<-filter(data,Substrate_Type=="S")  
```

## **Leaf/AMF**: Step-wise AIC Model Selection for 13C-PLFA
```{r leaf_AMF_variables, include=FALSE}
  leaf_AMF<-filter(leaf_PLFA,Microbial_Group=="AMF")
  leaf_AMF_13C<-leaf_AMF$d13
  Exclusion_AMF<-leaf_AMF$Rhizosphere_Manipulation
  Tree_AMF<-leaf_AMF$Tree
```

```{r leaf_AMF_stepAIC}
leaf_AMF_AIC<-stepAIC(object=aov(lm(leaf_AMF_13C~Exclusion_AMF*Tree_AMF)),direction="both",na.rm=TRUE)
```

```{r leaf_AMF_AIC_anova_table}
leaf_AMF_AIC$anova
```

```{r leaf_AMF_AIC_extraction}
print(extractAIC(aov(lm(leaf_AMF_13C~Exclusion_AMF+Tree_AMF))))
print(extractAIC(aov(lm(leaf_AMF_13C~Exclusion_AMF))))
```

Model selection indicates that **rhizosphere exclusion and tree species** should be included in model explaining **13C-PLFA** in **AMF** microbes inhabiting cores labeled with **leaf substrate**. Technically, best model includes on rhizosphere exclusion, but model that also includes tree species is included in the best family of models.

### Type II ANOVA Results
```{r leaf_AMF_ANOVA}
leafMod_AMF<-lm(d13~Rhizosphere_Manipulation+Tree,data=leaf_AMF)
Anova(leafMod_AMF,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r leaf_AMF_nested}
leaf_AMF_mixed_mod<-lme(d13~Rhizosphere_Manipulation+Tree,random=~1|Tree_ID,data=leaf_AMF,na.action=na.omit,method="ML")
leaf_AMF_fixed_mod<-lm(d13~Rhizosphere_Manipulation+Tree,data=leaf_AMF,na.action=na.omit)
anova(leaf_AMF_mixed_mod,leaf_AMF_fixed_mod)
```

Fixed model has lower AIC score than fixed model, with p-value of 0.95. Fixed model is clearly the superior model.

```{r leaf_AMF_fixed_mod}
summary(leaf_AMF_fixed_mod)
```

## **Leaf/Fungi**: Step-wise AIC Model Selection for 13C-PLFA
```{r leaf_fungi_variables, include=FALSE}
  leaf_fungi<-filter(leaf_PLFA,Microbial_Group=="FUNGI")
  leaf_fungi_13C<-leaf_fungi$d13
  Exclusion_fungi<-leaf_fungi$Rhizosphere_Manipulation
  Tree_fungi<-leaf_fungi$Tree
```

```{r leaf_fungi_stepAIC}
leaf_fungi_AIC<-stepAIC(object=aov(lm(leaf_fungi_13C~Exclusion_fungi*Tree_fungi)),direction="both",na.rm=TRUE)
```

```{r leaf_fungi_AIC_anova_table}
leaf_fungi_AIC$anova
```

```{r leaf_fungi_AIC_extraction}
print(extractAIC(aov(lm(leaf_fungi_13C~Exclusion_fungi+Tree_fungi))))
print(extractAIC(aov(lm(leaf_fungi_13C~Exclusion_fungi))))
```

Model selection indicates that **rhizosphere exclusion and tree species** should be included in model explaining **13C-PLFA** in **fungi** microbes inhabiting cores labeled with **leaf substrate**. Technically, best model includes on rhizosphere exclusion, but model that also includes tree species is included in the best family of models.

### Type II ANOVA Results
```{r leaf_fungi_ANOVA}
leafMod_fungi<-lm(d13~Rhizosphere_Manipulation+Tree,data=leaf_fungi)
Anova(leafMod_fungi,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r leaf_nested_fungi}
leaf_fungi_mixed_mod<-lme(d13~Rhizosphere_Manipulation+Tree,random=~1|Tree_ID,data=leaf_fungi,na.action=na.omit,method="ML")
leaf_fungi_fixed_mod<-lm(d13~Rhizosphere_Manipulation+Tree,data=leaf_fungi,na.action=na.omit)
anova(leaf_fungi_mixed_mod,leaf_fungi_fixed_mod)
```

Fixed model AIC score is ~2 points lower than mixed model AIC score, with p-value of 0.971, indicating that the fixed effects model is superior.

```{r leaf_fungi_fixed_mod}
summary(leaf_fungi_fixed_mod)
```

## **Leaf/Bacteria**: Step-wise AIC Model Selection for 13C-PLFA
```{r leaf_bact_variables, include=FALSE}
  leaf_bact<-filter(leaf_PLFA,Microbial_Group=="BACTERIA")
  leaf_bact_13C<-leaf_bact$d13
  Exclusion_bact<-leaf_bact$Rhizosphere_Manipulation
  Tree_bact<-leaf_bact$Tree
```

```{r leaf_bact_stepAIC}
leaf_bact_AIC<-stepAIC(object=aov(lm(leaf_bact_13C~Exclusion_bact*Tree_bact)),direction="both",na.rm=TRUE)
```

```{r leaf_bact_AIC_anova_table}
leaf_bact_AIC$anova
```

```{r leaf_bact_AIC_extraction}
print(extractAIC(aov(lm(leaf_bact_13C~Exclusion_bact+Tree_bact))))
print(extractAIC(aov(lm(leaf_bact_13C~Exclusion_bact))))
```

Model selection indicates that **rhizosphere exclusion and tree species** should be included in model explaining **13C-PLFA** in **bacteria** microbes inhabiting cores labeled with **leaf substrate**. Technically, best model includes on rhizosphere exclusion, but model that also includes tree species is included in the best family of models.

### Type II ANOVA Results
```{r leaf_bact_ANOVA}
leafMod_bact<-lm(d13~Rhizosphere_Manipulation+Tree,data=leaf_bact)
Anova(leafMod_bact,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r leaf_nested_bact}
leaf_bact_mixed_mod<-lme(d13~Rhizosphere_Manipulation+Tree,random=~1|Tree_ID,data=leaf_bact,na.action=na.omit,method="ML")
leaf_bact_fixed_mod<-lm(d13~Rhizosphere_Manipulation+Tree,data=leaf_bact,na.action=na.omit)
anova(leaf_bact_mixed_mod,leaf_bact_fixed_mod)
```

Fixed model AIC score is ~2 points lower than mixed model AIC score, with p-value of 0.551, indicating that the fixed effects model is superior.

```{r leaf_bact_fixed_mod}
summary(leaf_bact_fixed_mod)
```

## **Leaf/All Microbes**: Step-wise AIC Model Selection for 13C-PLFA
```{r leaf_allmicro_variables, include=FALSE}
  leaf_allmicro<-filter(leaf_PLFA,Microbial_Group=="ALL_MICROBES")
  leaf_allmicro_13C<-leaf_allmicro$d13
  Exclusion_allmicro<-leaf_allmicro$Rhizosphere_Manipulation
  Tree_allmicro<-leaf_allmicro$Tree
```

```{r leaf_allmicro_stepAIC}
leaf_allmicro_AIC<-stepAIC(object=aov(lm(leaf_allmicro_13C~Exclusion_allmicro*Tree_allmicro)),direction="both",na.rm=TRUE)
```

```{r leaf_allmicro_AIC_anova_table}
leaf_allmicro_AIC$anova
```

```{r leaf_allmicro_AIC_extraction}
print(extractAIC(aov(lm(leaf_allmicro_13C~Exclusion_allmicro+Tree_allmicro))))
print(extractAIC(aov(lm(leaf_allmicro_13C~Exclusion_allmicro))))
```

Model selection indicates that **rhizosphere exclusion and tree species** should be included in model explaining **13C-PLFA** in **all microbes** inhabiting cores labeled with **leaf substrate**. Technically, best model includes on rhizosphere exclusion, but model that also includes tree species is included in the best family of models.

### Type II ANOVA Results
```{r leaf_allmicro_ANOVA}
leafMod_allmicro<-lm(d13~Rhizosphere_Manipulation+Tree,data=leaf_allmicro)
Anova(leafMod_allmicro,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r leaf_nested_allmicro}
leaf_allmicro_mixed_mod<-lme(d13~Rhizosphere_Manipulation+Tree,random=~1|Tree_ID,data=leaf_allmicro,na.action=na.omit,method="ML")
leaf_allmicro_fixed_mod<-lm(d13~Rhizosphere_Manipulation+Tree,data=leaf_allmicro,na.action=na.omit)
anova(leaf_allmicro_mixed_mod,leaf_allmicro_fixed_mod)
```

Fixed model AIC score is ~2 points lower than mixed model AIC score, with p-value of 0.688, indicating that the fixed effects model is superior.

```{r leaf_allmicro_fixed_mod}
summary(leaf_allmicro_fixed_mod)
```

## **Starch/AMF**: Step-wise AIC Model Selection for 13C-PLFA
```{r starch_AMF_variables, include=FALSE}
  starch_AMF<-filter(starch_PLFA,Microbial_Group=="AMF")
  starch_AMF_13C<-starch_AMF$d13
  Exclusion_AMF<-starch_AMF$Rhizosphere_Manipulation
  Tree_AMF<-starch_AMF$Tree
```

```{r starch_AMF_stepAIC}
starch_AMF_AIC<-stepAIC(object=aov(lm(starch_AMF_13C~Exclusion_AMF*Tree_AMF)),direction="both",na.rm=TRUE)
```

```{r leaf_AMF_13C_PLFA_AIC_anova_table}
starch_AMF_AIC$anova
```

```{r leaf_AMF_13C_PLFA_AIC_extraction}
print(extractAIC(aov(lm(starch_AMF_13C~Exclusion_AMF+Tree_AMF))))
print(extractAIC(aov(lm(starch_AMF_13C~Exclusion_AMF))))
```

Model selection indicates that **rhizosphere exclusion and tree species** should be included in model explaining **13C-PLFA** in **AMF** microbes inhabiting cores labeled with **starch substrate**. Technically, best model includes on rhizosphere exclusion, but model that also includes tree species is included in the best family of models.

### Type II ANOVA Results
```{r starch_AMF_ANOVA}
starchMod_AMF<-lm(d13~Rhizosphere_Manipulation+Tree,data=starch_AMF)
Anova(starchMod_AMF,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r starch_AMF_nested}
starch_AMF_mixed_mod<-lme(d13~Rhizosphere_Manipulation+Tree,random=~1|Tree_ID,data=starch_AMF,na.action=na.omit,method="ML")
starch_AMF_fixed_mod<-lm(d13~Rhizosphere_Manipulation+Tree,data=starch_AMF,na.action=na.omit)
anova(starch_AMF_mixed_mod,starch_AMF_fixed_mod)
```

Fixed model has lower AIC score than fixed model, with p-value of 0.783. Fixed model is clearly the superior model.

```{r starch_AMF_fixed_mod}
summary(starch_AMF_fixed_mod)
```

## **Starch/Fungi**: Step-wise AIC Model Selection for 13C-PLFA
```{r starch_fungi_variables, include=FALSE}
  starch_fungi<-filter(starch_PLFA,Microbial_Group=="FUNGI")
  starch_fungi_13C<-starch_fungi$d13
  Exclusion_fungi<-starch_fungi$Rhizosphere_Manipulation
  Tree_fungi<-starch_fungi$Tree
```

```{r starch_fungi_stepAIC}
starch_fungi_AIC<-stepAIC(object=aov(lm(starch_fungi_13C~Exclusion_fungi*Tree_fungi)),direction="both",na.rm=TRUE)
```

```{r starch_fungi_AIC_anova_table}
starch_fungi_AIC$anova
```

```{r starch_fungi_AIC_extraction}
print(extractAIC(aov(lm(starch_fungi_13C~Exclusion_fungi+Tree_fungi))))
print(extractAIC(aov(lm(starch_fungi_13C~Exclusion_fungi))))
```

Model selection indicates that **rhizosphere exclusion and tree species** should be included in model explaining **13C-PLFA** in **fungi** microbes inhabiting cores labeled with **starch substrate**. Technically, best model includes on rhizosphere exclusion, but model that also includes tree species is included in the best family of models.

### Type II ANOVA Results
```{r starch_fungi_ANOVA}
starchMod_fungi<-lm(d13~Rhizosphere_Manipulation+Tree,data=starch_fungi)
Anova(starchMod_fungi,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r starch_nested_fungi}
starch_fungi_mixed_mod<-lme(d13~Rhizosphere_Manipulation+Tree,random=~1|Tree_ID,data=starch_fungi,na.action=na.omit,method="ML")
starch_fungi_fixed_mod<-lm(d13~Rhizosphere_Manipulation+Tree,data=starch_fungi,na.action=na.omit)
anova(starch_fungi_mixed_mod,starch_fungi_fixed_mod)
```

Fixed model AIC score is ~2 points lower than mixed model AIC score, with p-value of 0.9998, indicating that the fixed effects model is superior.

```{r starch_fungi_fixed_mod}
summary(starch_fungi_fixed_mod)
```

## **Starch/Bacteria**: Step-wise AIC Model Selection for 13C-PLFA
```{r starch_bact_variables, include=FALSE}
  starch_bact<-filter(starch_PLFA,Microbial_Group=="BACTERIA")
  starch_bact_13C<-starch_bact$d13
  Exclusion_bact<-starch_bact$Rhizosphere_Manipulation
  Tree_bact<-starch_bact$Tree
```

```{r starch_bact_stepAIC}
starch_bact_AIC<-stepAIC(object=aov(lm(starch_bact_13C~Exclusion_bact*Tree_bact)),direction="both",na.rm=TRUE)
```

```{r starch_bact_AIC_anova_table}
starch_bact_AIC$anova
```

```{r starch_bact_AIC_extraction}
print(extractAIC(aov(lm(starch_bact_13C~Exclusion_bact+Tree_bact))))
print(extractAIC(aov(lm(starch_bact_13C~Exclusion_bact))))
```

Model selection indicates that **rhizosphere exclusion and tree species** should be included in model explaining **13C-PLFA** in **bacteria** microbes inhabiting cores labeled with **starch substrate**. Technically, best model includes on rhizosphere exclusion, but model that also includes tree species is included in the best family of models.

### Type II ANOVA Results
```{r starch_bact_ANOVA}
starchMod_bact<-lm(d13~Rhizosphere_Manipulation+Tree,data=starch_bact)
Anova(starchMod_bact,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r starch_nested_bact}
starch_bact_mixed_mod<-lme(d13~Rhizosphere_Manipulation+Tree,random=~1|Tree_ID,data=starch_bact,na.action=na.omit,method="ML")
starch_bact_fixed_mod<-lm(d13~Rhizosphere_Manipulation+Tree,data=starch_bact,na.action=na.omit)
anova(starch_bact_mixed_mod,starch_bact_fixed_mod)
```

Fixed model AIC score is ~2 points lower than mixed model AIC score, with p-value of 0.797, indicating that the fixed effects model is superior.

```{r starch_bact_fixed_mod}
summary(starch_bact_fixed_mod)
```

## **Starch/All Microbes**: Step-wise AIC Model Selection for 13C-PLFA
```{r starch_allmicro_variables, include=FALSE}
  starch_allmicro<-filter(starch_PLFA,Microbial_Group=="ALL_MICROBES")
  starch_allmicro_13C<-starch_allmicro$d13
  Exclusion_allmicro<-starch_allmicro$Rhizosphere_Manipulation
  Tree_allmicro<-starch_allmicro$Tree
```

```{r starch_allmicro_stepAIC}
starch_allmicro_AIC<-stepAIC(object=aov(lm(starch_allmicro_13C~Exclusion_allmicro*Tree_allmicro)),direction="both",na.rm=TRUE)
```

```{r starch_allmicro_AIC_anova_table}
starch_allmicro_AIC$anova
```

```{r starch_allmicro_AIC_extraction}
print(extractAIC(aov(lm(starch_allmicro_13C~Exclusion_allmicro+Tree_allmicro))))
print(extractAIC(aov(lm(starch_allmicro_13C~Exclusion_allmicro))))
```

Model selection indicates that **rhizosphere exclusion and tree species** should be included in model explaining **13C-PLFA** in **all microbes** inhabiting cores labeled with **starch substrate**. Technically, best model includes on rhizosphere exclusion, but model that also includes tree species is included in the best family of models.

### Type II ANOVA Results
```{r starch_allmicro_ANOVA}
starchMod_allmicro<-lm(d13~Rhizosphere_Manipulation+Tree,data=starch_allmicro)
Anova(starchMod_allmicro,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r starch_nested_allmicro}
starch_allmicro_mixed_mod<-lme(d13~Rhizosphere_Manipulation+Tree,random=~1|Tree_ID,data=starch_allmicro,na.action=na.omit,method="ML")
starch_allmicro_fixed_mod<-lm(d13~Rhizosphere_Manipulation+Tree,data=starch_allmicro,na.action=na.omit)
anova(starch_allmicro_mixed_mod,starch_allmicro_fixed_mod)
```

Fixed model AIC score is ~2 points lower than mixed model AIC score, with p-value of 0.9998, indicating that the fixed effects model is superior.

```{r starch_allmicro_fixed_mod}
summary(starch_allmicro_fixed_mod)
```

# 13C Specific Respiration Models
```{r read_13C_specResp_data, include=FALSE}
  ## Read in data
  data<-read.csv("Final_Specific_Respiration_Data.csv")
  data$Mesh_Type<-as.factor(data$Mesh_Type)
  data$Tree_ID<-substring(data$Core_ID,1,5)
  leaf<-filter(data,Label_Type=="L")
  starch<-filter(data,Label_Type=="S")
  
```

## **Leaf Substrate**: Step-wise AIC Model Selection for 13C Specific Respiration
```{r leaf_13C_specResp_variables, include=FALSE}
  leaf_13C_specresp<-leaf$Specific_Respiration_CO2_FAME
  Exclusion<-leaf$Mesh_Type
  Tree<-leaf$Tree_Species
```

```{r leaf_13C_specResp_stepAIC}
leaf_13C_specresp_AIC<-stepAIC(object=aov(lm(leaf_13C_specresp~Exclusion*Tree+Exclusion+Tree)),direction="both",na.rm=TRUE)
```

```{r leaf_13C_specResp_AIC_anova_table}
leaf_13C_specresp_AIC$anova
```

```{r leaf_13C_specResp_AIC_extraction}
print(extractAIC(aov(lm(leaf_13C_specresp~Exclusion*Tree))))
```

Model selection indicates that all parameters, i.e. **rhizosphere exclusion, tree species, and exclusion x tree interaction** should be included in model explaining **13C-specific respiration** in cores labeled with **leaf substrate**. 

### Type II ANOVA Results
```{r leaf_13C_specResp_ANOVA}
leaf_mod<-lm(Specific_Respiration_CO2_FAME~Mesh_Type*Tree_Species,data=leaf)
Anova(leaf_mod,type="II")
```

### Tukey's HSD Results
```{r leaf_13C_specResp_TukeyHSD}
tx<-with(leaf, interaction(Mesh_Type,Tree_Species))
amod<-aov(Specific_Respiration_CO2_FAME ~ tx, data=leaf)
HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

### Compare Nested Model to Fixed Effects Model
```{r leaf_nested_BSR}
leaf_BSR_mixed_mod<-lme(Specific_Respiration_CO2_FAME~Mesh_Type*Tree_Species,random=~1|Tree_ID,data=leaf,na.action=na.omit,method="ML")
leaf_BSR_fixed_mod<-lm(Specific_Respiration_CO2_FAME~Mesh_Type*Tree_Species,data=leaf,na.action=na.omit)
anova(leaf_BSR_mixed_mod,leaf_BSR_fixed_mod)
```

This output indicates that the AIC score for the mixed model is ~2 points higher, mixed model does not perform better.

```{r leaf_BSR_fixed_mod}
summary(leaf_BSR_fixed_mod)
```

## **Starch Substrate**: Step-wise AIC Model Selection for 13C Specific Respiration
```{r starch_13C_specResp_variables, include=FALSE}
  starch_13C_specresp<-starch$Specific_Respiration_CO2_FAME
  Exclusion<-starch$Mesh_Type
  Tree<-starch$Tree_Species
```

```{r starch_13C_specResp_stepAIC}
starch_13C_specresp_AIC<-stepAIC(object=aov(lm(starch_13C_specresp~Exclusion+Tree+Exclusion*Tree)),direction="both",na.rm=TRUE)
```

```{r starch_13C_specResp_AIC_anova_table}
starch_13C_specresp_AIC$anova
```

```{r starch_13C_specResp_AIC_extraction}
print(extractAIC(aov(lm(starch_13C_specresp~Exclusion+Tree))))
print(extractAIC(aov(lm(starch_13C_specresp~Exclusion))))
print(extractAIC(aov(lm(starch_13C_specresp~Tree))))
print(extractAIC(aov(lm(starch_13C_specresp~1))))
```

Model selection indicates that the best model for **13C-specific respiration of starch substrate** includes **only the intercept**, but the model including **tree species** as a parameter is within the family of best models. 

### Type II ANOVA Results
```{r starch_13C_specResp_ANOVA}
starch_mod<-lm(Specific_Respiration_CO2_FAME~Mesh_Type+Tree_Species,data=starch)
Anova(starch_mod,type="II")
```

### Tukey's HSD Results
```{r starch_13C_specResp_TukeyHSD}
tx<-with(starch, interaction(Mesh_Type,Tree_Species))
amod<-aov(Specific_Respiration_CO2_FAME ~ tx, data=starch)
HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

### Compare Nested Model to Fixed Effects Model
```{r starch_nested_BSR}
starch_BSR_mixed_mod<-lme(Specific_Respiration_CO2_FAME~Mesh_Type*Tree_Species,random=~1|Tree_ID,data=starch,na.action=na.omit,method="ML")
starch_BSR_fixed_mod<-lm(Specific_Respiration_CO2_FAME~Mesh_Type*Tree_Species,data=starch,na.action=na.omit)
anova(starch_BSR_mixed_mod,starch_BSR_fixed_mod)
```

This output indicates that the AIC score for the mixed model is ~2 points higher, mixed model does not perform better.

```{r starch_BSR_fixed_mod}
summary(starch_BSR_fixed_mod)
```


# Background Respiration Models

## Step-wise AIC Model Selection for Background Respiration
```{r read_back_resp_data, include=FALSE}
data<-read.csv("Resp_Data_for_Spec_Resp.csv")
data$Tree_ID<-substring(data$core_ID,1,5)
```

```{r back_resp_variables}
Resp<-data$Mean_Respiration
Tree<-data$Tree_species
Exclusion<-data$Exclusion
```

```{r back_resp_stepAIC}
back_resp_AIC<-stepAIC(object=aov(lm(Resp~Exclusion+Tree+Exclusion*Tree)),direction="both",na.rm=TRUE)
```

```{r back_resp_AIC_anova_table}
back_resp_AIC$anova
```

```{r back_resp_AIC_extraction}
print(extractAIC(aov(lm(Resp~Exclusion))))
print(extractAIC(aov(lm(Resp~Tree))))
print(extractAIC(aov(lm(Resp~Exclusion+Tree))))
```

Model selection indicates that the best model for **background respiration** includes **rhizosphere exclusion and tree species**. Model containing only rhizosphere exclusion is technically the best model, but the model containing both rhizosphere exclusion and tree species falls within the "family of best models".

### Type II ANOVA Results
```{r back_resp_ANOVA}
resp_mod<-lm(Mean_Respiration~Exclusion+Tree_species,data=data)
Anova(resp_mod,type="II")
```

### Tukey's HSD Results
```{r back_resp_TukeyHSD}
tx<-with(data, interaction(Exclusion,Tree_species))
amod<-aov(Mean_Respiration ~ tx, data=data)
HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

```{r back_resp_TukeyHSD_Exclusion}
amod<-aov(data$Mean_Respiration~as.factor(data$Exclusion))
TukeyHSD(amod)
```

### Compare Nested Model to Fixed Effects Model
```{r back_resp_nested}
resp_mixed_mod<-lme(Mean_Respiration~Exclusion+Tree_species,random=~1|Tree_ID,data=data,na.action=na.omit,method="ML")
resp_fixed_mod<-lm(Mean_Respiration~Exclusion+Tree_species,data=data,na.action=na.omit)
anova(resp_mixed_mod,resp_fixed_mod)
```

AIC output for fixed and mixed models practically identical; random effect does not improve model.

```{r back_resp_fixed_mod}
summary(resp_fixed_mod)
```

# Background AMF FAME Models

## Step-wise AIC Model Selection for Background AMF
```{r read_back_PLFA_data, include=FALSE}
data<-read.csv("Unlab_PLFA_for_Spec_Resp.csv")
data$Tree_ID<-substring(data$Sample_Name,1,5)
data$Rhizosphere_Manipulation<-as.factor(data$Rhizosphere_Manipulation)
```

```{r back_AMF_variables, include=FALSE}
AMF_PLFA<-filter(data,Microbial_Group=="AMF")
FAME_AMF<-AMF_PLFA$ln_FAME
Tree_AMF<-AMF_PLFA$Tree
Exclusion_AMF<-AMF_PLFA$Rhizosphere_Manipulation
```

```{r back_AMF_stepAIC}
back_AMF_AIC<-stepAIC(object=aov(lm(FAME_AMF~Exclusion_AMF*Tree_AMF)),direction="both",na.rm=TRUE)
```

```{r back_AMF_AIC_anova_table}
back_AMF_AIC$anova
```

```{r back_AMF_AIC_extraction}
print(extractAIC(aov(lm(FAME_AMF~Exclusion_AMF+Tree_AMF))))
print(extractAIC(aov(lm(FAME_AMF~Exclusion_AMF))))
print(extractAIC(aov(lm(FAME_AMF~Tree_AMF))))
```

Model selection indicates that the best model for **background AMF FAME** includes **rhizosphere exclusion and tree species**, but no interaction terms. Technically, the model that includes only tree species as a fixed effect is the best model, but the model including tree species and rhizosphere exclusion as fixed effects is only ~2 AIC points higher, and is therefore included in the family of best models.

### Type II ANOVA Results
```{r back_AMF_ANOVA}
back_AMF_mod<-lm(ln_FAME~Rhizosphere_Manipulation+Tree,data=AMF_PLFA)
Anova(back_AMF_mod,type="II")
```

### Compare Nested Model to Fixed Effects Model
```{r back_AMF_nested}
back_AMF_mixed_mod<-lme(ln_FAME~Rhizosphere_Manipulation+Tree,random=~1|Tree_ID,data=AMF_PLFA,na.action=na.omit,method="ML")
back_AMF_fixed_mod<-lm(ln_FAME~Rhizosphere_Manipulation+Tree,data=AMF_PLFA,na.action=na.omit)
anova(back_AMF_mixed_mod,back_AMF_fixed_mod)
```

Mixed effects model is ~3 AIC points lower than the fixed effects model, with a p-value of 0.019, so the **mixed effects model is superior** and there is significant variation attributed to individual trees across the field site.

```{r back_AMF_mixed_mod}
summary(back_AMF_mixed_mod)
```

# Background Fungi FAME Models

## Step-wise AIC Model Selection for Background Fungi
```{r back_fungi_variables, include=FALSE}
fungi_PLFA<-filter(data,Microbial_Group=="FUNGI")
FAME_fungi<-fungi_PLFA$ln_FAME
Tree_fungi<-fungi_PLFA$Tree
Exclusion_fungi<-fungi_PLFA$Rhizosphere_Manipulation
```

```{r back_fungi_stepAIC}
back_fungi_AIC<-stepAIC(object=aov(lm(FAME_fungi~Exclusion_fungi*Tree_fungi)),direction="both",na.rm=TRUE)
```

```{r back_fungi_AIC_anova_table}
back_fungi_AIC$anova
```

```{r back_fungi_AIC_extraction}
print(extractAIC(aov(lm(FAME_fungi~Exclusion_fungi+Tree_fungi))))
print(extractAIC(aov(lm(FAME_fungi~Exclusion_fungi))))
print(extractAIC(aov(lm(FAME_fungi~Tree_fungi))))
```

Model selection indicates that the best model for **background fungi FAME** includes **just tree species**, and no interaction terms. The model that includes both tree species and rhizosphere exclusion as fixed effects is ~4 AIC points higher, indicating that the model with only tree species as a fixed effect is superior.

### Type II ANOVA Results
```{r back_fungi_ttest}
t.test(ln_FAME~Tree,data=fungi_PLFA,alternative="two.sided")
```

### Compare Nested Model to Fixed Effects Model
```{r back_fungi_nested}
back_fungi_mixed_mod<-lme(ln_FAME~Tree,random=~1|Tree_ID,data=fungi_PLFA,na.action=na.omit,method="ML")
back_fungi_fixed_mod<-lm(ln_FAME~Tree,data=fungi_PLFA,na.action=na.omit)
anova(back_fungi_mixed_mod,back_fungi_fixed_mod)
```

Fixed effects model is ~1.5 AIC points lower than the mixed effects model, with a p-value of 0.4371, so the **fixed effects model is superior**.

```{r back_fungi_fixed_mod}
summary(back_fungi_fixed_mod) # T-test results from above are the more appropriate test.
```

# Background Bacteria FAME Models

## Step-wise AIC Model Selection for Background Bacteria
```{r back_bact_variables, include=FALSE}
bact_PLFA<-filter(data,Microbial_Group=="BACTERIA")
FAME_bact<-bact_PLFA$ln_FAME
Tree_bact<-bact_PLFA$Tree
Exclusion_bact<-bact_PLFA$Rhizosphere_Manipulation
```

```{r back_bact_stepAIC}
back_bact_AIC<-stepAIC(object=aov(lm(FAME_bact~Exclusion_bact*Tree_bact)),direction="both",na.rm=TRUE)
```

```{r back_bact_AIC_anova_table}
back_bact_AIC$anova
```

```{r back_bact_AIC_extraction}
print(extractAIC(aov(lm(FAME_bact~Exclusion_bact+Tree_bact))))
print(extractAIC(aov(lm(FAME_bact~Exclusion_bact))))
print(extractAIC(aov(lm(FAME_bact~Tree_bact))))
```

Model selection indicates that the best model for **background bacteria FAME** includes **rhizosphere exclusion and tree species**, but no interaction terms. Technically, the model that includes only tree species as a fixed effect is the best model, but the model including tree species and rhizosphere exclusion as fixed effects is only ~1.5 AIC points higher, and is therefore included in the family of best models.

### Type II ANOVA Results
```{r back_bact_ANOVA}
back_bact_mod<-lm(ln_FAME~Rhizosphere_Manipulation+Tree,data=bact_PLFA)
Anova(back_bact_mod,type="II")
```

### Compare Nested Model to Fixed Effects Model
```{r back_bact_nested}
back_bact_mixed_mod<-lme(ln_FAME~Rhizosphere_Manipulation+Tree,random=~1|Tree_ID,data=bact_PLFA,na.action=na.omit,method="ML")
back_bact_fixed_mod<-lm(ln_FAME~Rhizosphere_Manipulation+Tree,data=bact_PLFA,na.action=na.omit)
anova(back_bact_mixed_mod,back_bact_fixed_mod)
```

These models are very close in there AIC scores; the mixed effects model is ~0.9 AIC points lower than the fixed effects model, with a p-value of 0.090, meaning that either model is valid. I will therefore select the more simple, fixed effects model as the superior model.

```{r back_bact_fixed_mod}
summary(back_bact_fixed_mod)
```

# Background All Microbes FAME Models

## Step-wise AIC Model Selection for Background FAME Abundance of All Microbes
```{r back_allmicro_variables, include=FALSE}
allmicro_PLFA<-filter(data,Microbial_Group=="ALL_MICROBES")
FAME_allmicro<-allmicro_PLFA$ln_FAME
Tree_allmicro<-allmicro_PLFA$Tree
Exclusion_allmicro<-allmicro_PLFA$Rhizosphere_Manipulation
```

```{r back_allmicro_stepAIC}
back_allmicro_AIC<-stepAIC(object=aov(lm(FAME_allmicro~Exclusion_allmicro*Tree_allmicro)),direction="both",na.rm=TRUE)
```

```{r back_allmicro_AIC_anova_table}
back_allmicro_AIC$anova
```

```{r back_allmicro_AIC_extraction}
print(extractAIC(aov(lm(FAME_allmicro~Exclusion_allmicro+Tree_allmicro))))
print(extractAIC(aov(lm(FAME_allmicro~Exclusion_allmicro))))
print(extractAIC(aov(lm(FAME_allmicro~Tree_allmicro))))
```

Model selection indicates that the best model for **background FAME for all microbes** includes **rhizosphere exclusion and tree species**, but no interaction terms. Technically, the model that includes only tree species as a fixed effect is the best model, but the model including tree species and rhizosphere exclusion as fixed effects is only ~2 AIC points higher, and is therefore included in the family of best models.

### Type II ANOVA Results
```{r back_allmicro_ANOVA}
back_allmicro_mod<-lm(ln_FAME~Rhizosphere_Manipulation+Tree,data=allmicro_PLFA)
Anova(back_allmicro_mod,type="II")
```

### Compare Nested Model to Fixed Effects Model
```{r back_allmicro_nested}
back_allmicro_mixed_mod<-lme(ln_FAME~Rhizosphere_Manipulation+Tree,random=~1|Tree_ID,data=allmicro_PLFA,na.action=na.omit,method="ML")
back_allmicro_fixed_mod<-lm(ln_FAME~Rhizosphere_Manipulation+Tree,data=allmicro_PLFA,na.action=na.omit)
anova(back_allmicro_mixed_mod,back_allmicro_fixed_mod)
```

These models are very close in there AIC scores; the mixed effects model is ~0.8 AIC points lower than the fixed effects model, with a p-value of 0.099, meaning that either model is valid. I will therefore select the more simple, fixed effects model as the superior model.

```{r back_allmicro_fixed_mod}
summary(back_allmicro_fixed_mod)
```

# Background Specific Respiration Models

## Step-wise AIC Model Selection for Background Specific Respiration
```{r read_back_specresp_data, include=FALSE}
data<-read.csv("Unlab_Spec_Resp_CO2_and_PLFA_Data.csv")
```

```{r back_specresp_variables}
specResp<-data$Specific_Respiration_CO2_FAME
Exclusion<-data$Rhizosphere_Manipulation
Tree<-data$Tree
```

```{r back_specresp_stepAIC}
back_specresp_AIC<-stepAIC(object=aov(lm(specResp~Exclusion*Tree)),direction="both",na.rm=TRUE)
```

```{r back_specresp_AIC_anova_table}
back_specresp_AIC$anova
```

Model selection indicates that the best model for **background specific respiration** includes **rhizosphere exclusion and tree species** (no interaction term).

### Type II ANOVA Results
```{r back_specresp_ANOVA}
back_specresp_mod<-lm(specResp~Exclusion+Tree,data=data)
Anova(back_specresp_mod,type="II")
```

### Tukey's HSD Results
```{r back_specresp_TukeyHSD}
tx<-with(data, interaction(Exclusion,Tree))
amod<-aov(specResp ~ tx, data=data)
HSD.test(amod, "tx", group=TRUE, console=TRUE)
```
