---
title: "Costa Rica Carbon Stats Output"
author: "Kenna Rewcastle"
date: "4/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(dplyr)
library(MASS)
library(car)
library(agricolae)
library(tidyr)
library(reshape2)
```


# Delta 13C-CO2 Models

```{r, include=FALSE}
  ## Read in data
  dat<-read.csv("Costa Rica Master sheet annotated.csv")
  dat<-filter(dat,dat$Exclusion!="NA")
  dat$Exclusion<-as.factor(dat$Exclusion)

  ## Separate starch and leaf data
  starch_CO2<-filter(dat,dat$Isotope_label=="S") 
  leaf_CO2<-filter(dat,dat$Isotope_label=="L")
```

## **Leaf Substrate**: Step-wise AIC Model Selection for 13C-CO2
```{r, include=FALSE}
    leaf_d13<-leaf_CO2$d13_d9
    Exclusion<-leaf_CO2$Exclusion
    Tree_Species<-leaf_CO2$Tree_species
```

```{r}
leaf_CO2_AIC<-stepAIC(object=aov(lm(leaf_d13~Exclusion+Tree_Species+Exclusion*Tree_Species)),direction="both",na.rm=TRUE)
```

```{r}
leaf_CO2_AIC$anova
```

Model selection indicates that **rhizosphere exclusion, tree species, and exclusion x tree species** should all be included in model explaining **13C-CO2 efflux** from cores labeled with **leaf substrate**.

### Type II ANOVA Results
```{r}
leafMod<-lm(d13_d9~Exclusion*Tree_species,data=leaf_CO2)
  Anova(leafMod,type="II") 
```

### Tukey's HSD Results
```{r}
 tx<-with(leaf_CO2, interaction(Exclusion, Tree_species))
  amod<-aov(d13_d9 ~ tx, data=leaf_CO2)
  leaf_CO2_Tukey<-HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

## **Starch Substrate**: Step-wise AIC Model Selection for 13C-CO2
```{r, include=FALSE}
    starch_d13<-starch_CO2$d13_d2
    Exclusion<-starch_CO2$Exclusion
    Tree_Species<-starch_CO2$Tree_species
```

```{r}
starch_CO2_AIC<-stepAIC(object=aov(lm(starch_d13~Exclusion+Tree_Species+Exclusion*Tree_Species)),direction="both",na.rm=TRUE)
```

```{r}
starch_CO2_AIC$anova
```

```{r}
print(extractAIC(aov(lm(starch_d13~Exclusion))))
print(extractAIC(aov(lm(starch_d13~Exclusion+Tree_Species))))
print(extractAIC(aov(lm(starch_d13~Exclusion*Tree_Species))))
```

Model selection indicates that **rhizosphere exclusion and tree species** should be included in model explaining **13C-CO2 efflux** from cores labeled with **starch substrate**. Model with only rhizosphere exclusion and model with rhizosphere exclusion and tree species are separated by only 0.7 AIC points, so both models are included in "best family" of models. Will choose to include tree species as a parameter to match structure of leaf models.

### Type II ANOVA Results
```{r}
starchMod<-lm(d13_d2~Exclusion+Tree_species,data=starch_CO2)
  Anova(starchMod,type="II") 
```

### Tukey's HSD Results
```{r}
  tx<-with(starch_CO2, interaction(Exclusion, Tree_species))
  amod<-aov(d13_d2 ~ tx, data=starch_CO2)
  HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

# Delta 13C-PLFA Models
```{r, include=FALSE}
  ## Read in data
  data<-read.csv("FINAL_PLFA_with_metadata.csv")
  data<-data[,2:15] # Gets rid of ID column
  data<-filter(data,Microbial_Group!="standard")
  data$Rhizosphere_Manipulation<-as.factor(data$Rhizosphere_Manipulation)
  data$umol_FAME<-data$nmol_FAME_per_g_soil/1000
  
  ## Simplify the microbial groupings
  data_trim<-data.frame("Sample_Name"=data[,1],"Tree"=data[,2],"Rhizosphere_Manipulation"=data[,3],"Substrate_Type"=data[,4],"Microbial_Group"=data[,14],"d13"=data[,9])
  
  data_wide<-dcast(data_trim,Sample_Name+Tree+Rhizosphere_Manipulation+Substrate_Type ~ Microbial_Group, value.var="d13", fun.aggregate=mean, na.rm=TRUE)
  
  is.nan.data.frame <- function(x)
    do.call(cbind, lapply(x, is.nan))
  data_wide[is.nan(data_wide)] <- NA
  
      ### Take mean of narrow groups to yield FUNGI, BACTERIA, ALL_MICROBE groups
      for (i in 1:nrow(data_wide)) {
        data_wide$BACTERIA[i]<-mean(c(data_wide$bacteria[i],data_wide$`gram+bacteria`[i],data_wide$`gram-bacteria`[i]),na.rm=TRUE)
        data_wide$FUNGI[i]<-mean(c(data_wide$AMF[i],data_wide$fungi[i]),na.rm=TRUE)
        data_wide$ALL_MICROBES[i]<-mean(c(data_wide$AMF[i],data_wide$bacteria[i],data_wide$fungi[i],data_wide$general[i],data_wide$`gram-bacteria`[i],data_wide$`gram+bacteria`[i],data_wide$protozoa[i]),na.rm=TRUE)
      }

  data_wide<-data_wide[,c(1,2,3,4,5,12,13,14)] # Trim down to broader groups
  
  data<-melt(data_wide,value.name="d13",id.vars=c("Sample_Name","Rhizosphere_Manipulation","Tree","Substrate_Type"))
  names(data)[5]<-"Microbial_Group"
  
  ## Separate leaf and starch datasets
  leaf_PLFA<-filter(data,Substrate_Type=="L")
  starch_PLFA<-filter(data,Substrate_Type=="S")  
```

## **Starch Substrate**: Step-wise AIC Model Selection for 13C-PLFA
```{r, include=FALSE}
  starch_13C_PLFA<-starch_PLFA$d13
  Exclusion<-starch_PLFA$Rhizosphere_Manipulation
  Tree<-starch_PLFA$Tree
  Microbe_Group<-starch_PLFA$Microbial_Group
```

```{r}
starch_13CPLFA_AIC<-stepAIC(object=aov(lm(starch_13C_PLFA~Exclusion*Tree*Microbe_Group)),direction="both",na.rm=TRUE)
```

```{r}
starch_13CPLFA_AIC$anova
```

```{r}
print(extractAIC(aov(lm(starch_13C_PLFA~Exclusion+Microbe_Group))))
print(extractAIC(aov(lm(starch_13C_PLFA~Exclusion+Microbe_Group+Tree))))
```

Model selection indicates that **rhizosphere exclusion, microbial group, and tree species** should be included in model explaining **13C-PLFA** in cores labeled with **starch substrate**. Model with only rhizosphere exclusion and microbial group and model with rhizosphere exclusion, microbial group, and tree species are separated by only 1 AIC point, so both models are included in "best family" of models. Will choose to include tree species as a parameter to match structure of CO2 Analyses.

### Type II ANOVA Results
```{r}
starchMod_13CPLFA<-lm(d13~Rhizosphere_Manipulation+Microbial_Group+Tree,data=starch_PLFA)
Anova(starchMod_13CPLFA,type="II") 
```

### Tukey's HSD Results
```{r}
tx<-with(starch_PLFA, interaction(Rhizosphere_Manipulation,Microbial_Group,Tree))
amod<-aov(d13 ~ tx, data=starch_PLFA)
HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

## **Leaf Substrate**: Step-wise AIC Model Selection for 13C-PLFA
```{r, include=FALSE}
  leaf_13C_PLFA<-leaf_PLFA$d13
  Exclusion<-leaf_PLFA$Rhizosphere_Manipulation
  Tree<-leaf_PLFA$Tree
  Microbe_Group<-leaf_PLFA$Microbial_Group
```

```{r}
leaf_13CPLFA_AIC<-stepAIC(object=aov(lm(leaf_13C_PLFA~Exclusion*Tree*Microbe_Group)),direction="both",na.rm=TRUE)
```

```{r}
leaf_13CPLFA_AIC$anova
```

```{r}
print(extractAIC(aov(lm(leaf_13C_PLFA~Exclusion+Tree+Exclusion*Tree))))
print(extractAIC(aov(lm(leaf_13C_PLFA~Exclusion+Tree+Exclusion*Tree+Microbe_Group))))
```

Model selection indicates that **rhizosphere exclusion, tree species, and exclusion x tree interaction** should be included in model explaining **13C-PLFA** in cores labeled with **leaf substrate**. Model that included microbial group was ~6 AIC points greater than the final model, so microbial group was not included in final analyses.

### Type II ANOVA Results
```{r}
leafMod_13CPLFA<-lm(d13~Rhizosphere_Manipulation+Tree+Rhizosphere_Manipulation*Tree,data=leaf_PLFA)
Anova(leafMod_13CPLFA,type="II") 
```

### Tukey's HSD Results
```{r}
tx<-with(leaf_PLFA, interaction(Rhizosphere_Manipulation,Microbial_Group,Tree))
amod<-aov(d13 ~ tx, data=leaf_PLFA)
HSD.test(amod, "tx", group=TRUE, console=TRUE)
```
