---
title: "Costa Rica Carbon Stats Output"
author: "Kenna Rewcastle"
date: "4/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(dplyr)
library(MASS)
library(car)
library(agricolae)
library(tidyr)
library(reshape2)
library(nlme)
library(RLRsim)
```


# Delta 13C-CO2 Models

```{r read_13C_CO2_data, include=FALSE}
  ## Read in data
  dat<-read.csv("Costa Rica Master sheet annotated.csv")
  dat<-filter(dat,dat$Exclusion!="NA")
  dat$Exclusion<-as.factor(dat$Exclusion)
  dat$Tree_ID<-substr(dat$core_ID,1,5)

  ## Separate starch and leaf data
  starch_CO2<-filter(dat,dat$Isotope_label=="S") 
  leaf_CO2<-filter(dat,dat$Isotope_label=="L")
```

## **Leaf Substrate**: Step-wise AIC Model Selection for 13C-CO2
```{r leaf_13C_CO2_variables, include=FALSE}
    leaf_d13<-leaf_CO2$d13_d9
    Exclusion<-leaf_CO2$Exclusion
    Tree_Species<-leaf_CO2$Tree_species
```

```{r leaf_13C_CO2_stepAIC}
leaf_CO2_AIC<-stepAIC(object=aov(lm(leaf_d13~Exclusion+Tree_Species+Exclusion*Tree_Species)),direction="both",na.rm=TRUE)
```

```{r leaf_13C_CO2_AIC_anova_table}
leaf_CO2_AIC$anova
```

Model selection indicates that **rhizosphere exclusion, tree species, and exclusion x tree species** should all be included in model explaining **13C-CO2 efflux** from cores labeled with **leaf substrate**.

### Type II ANOVA Results
```{r leaf_13C_CO2_ANOVA}
leafMod<-lm(d13_d9~Exclusion*Tree_species,data=leaf_CO2,na.action=na.omit)
leaf_CO2_fixed<-Anova(leafMod,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r leaf_nested_13CO2}
leaf_CO2_mixed_mod<-lme(d13_d9~Exclusion*Tree_species,random=~1|Tree_ID,data=leaf_CO2,na.action=na.omit,method="ML")
leaf_CO2_fixed_mod<-lm(d13_d9~Exclusion*Tree_species,data=leaf_CO2,na.action=na.omit)
anova(leaf_CO2_mixed_mod,leaf_CO2_fixed_mod)
```

This output indicates that the AIC score for the mixed model is ~1.5 points lower than the fixed effects model, while BIC is virtually identical, and the p-value is 0.0639.TL;DR: Unclear whether tree ID random intercept is necessary. Evaluate other models first.

```{r leaf_CO2_mixed_mod}
summary(leaf_CO2_mixed_mod)
```

```{r leaf_CO2_fixed_mod}
summary(leaf_CO2_fixed_mod)
```

### Tukey's HSD Results
```{r leaf_13C_CO2_TukeyHSD}
 tx<-with(leaf_CO2, interaction(Exclusion, Tree_species))
  amod<-aov(d13_d9 ~ tx, data=leaf_CO2)
  leaf_CO2_Tukey<-HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

## **Starch Substrate**: Step-wise AIC Model Selection for 13C-CO2
```{r starch_13C_CO2_variables, include=FALSE}
    starch_d13<-starch_CO2$d13_d2
    Exclusion<-starch_CO2$Exclusion
    Tree_Species<-starch_CO2$Tree_species
```

```{r starch_13C_CO2_stepAIC}
starch_CO2_AIC<-stepAIC(object=aov(lm(starch_d13~Exclusion+Tree_Species+Exclusion*Tree_Species)),direction="both",na.rm=TRUE)
```

```{r starch_13C_CO2_AIC_anova_table}
starch_CO2_AIC$anova
```

```{r starch_13C_CO2_AIC_extraction}
print(extractAIC(aov(lm(starch_d13~Exclusion))))
print(extractAIC(aov(lm(starch_d13~Exclusion+Tree_Species))))
print(extractAIC(aov(lm(starch_d13~Exclusion*Tree_Species))))
```

Model selection indicates that **rhizosphere exclusion and tree species** should be included in model explaining **13C-CO2 efflux** from cores labeled with **starch substrate**. Model with only rhizosphere exclusion and model with rhizosphere exclusion and tree species are separated by only 0.7 AIC points, so both models are included in "best family" of models. Will choose to include tree species as a parameter to match structure of leaf models.

### Type II ANOVA Results
```{r starch_13C_CO2_ANOVA}
starchMod<-lm(d13_d2~Exclusion+Tree_species,data=starch_CO2)
  Anova(starchMod,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r starch_nested_13CO2}
starch_CO2_mixed_mod<-lme(d13_d2~Exclusion+Tree_species,random=~1|Tree_ID,data=starch_CO2,na.action=na.omit,method="ML")
starch_CO2_fixed_mod<-lm(d13_d2~Exclusion+Tree_species,data=starch_CO2,na.action=na.omit)
anova(starch_CO2_mixed_mod,starch_CO2_fixed_mod)
```

This output indicates that the AIC score for the mixed model is ~2.6 points lower than the fixed effects model, while BIC scores are only 1 point apart, and the p-value is 0.0305.TL;DR: Slightly more decisive that the mixed effects model is best.

```{r starch_CO2_mixed_mod}
summary(starch_CO2_mixed_mod)
```

```{r starch_CO2_fixed_mod}
summary(starch_CO2_fixed_mod)
```


### Tukey's HSD Results
```{r starch_13C_CO2_TukeyHSD}
  tx<-with(starch_CO2, interaction(Exclusion, Tree_species))
  amod<-aov(d13_d2 ~ tx, data=starch_CO2)
  HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

# Delta 13C-PLFA Models
```{r read_13C_PLFA_data, include=FALSE}
  ## Read in data
  data<-read.csv("FINAL_PLFA_with_metadata.csv")
  data<-data[,2:15] # Gets rid of ID column
  data<-filter(data,Microbial_Group!="standard")
  data$Rhizosphere_Manipulation<-as.factor(data$Rhizosphere_Manipulation)
  data$umol_FAME<-data$nmol_FAME_per_g_soil/1000
  
  ## Simplify the microbial groupings
  data_trim<-data.frame("Sample_Name"=data[,1],"Tree"=data[,2],"Rhizosphere_Manipulation"=data[,3],"Substrate_Type"=data[,4],"Microbial_Group"=data[,14],"d13"=data[,9])
  
  data_wide<-dcast(data_trim,Sample_Name+Tree+Rhizosphere_Manipulation+Substrate_Type ~ Microbial_Group, value.var="d13", fun.aggregate=mean, na.rm=TRUE)
  
  is.nan.data.frame <- function(x)
    do.call(cbind, lapply(x, is.nan))
  data_wide[is.nan(data_wide)] <- NA
  
      ### Take mean of narrow groups to yield FUNGI, BACTERIA, ALL_MICROBE groups
      for (i in 1:nrow(data_wide)) {
        data_wide$BACTERIA[i]<-mean(c(data_wide$bacteria[i],data_wide$`gram+bacteria`[i],data_wide$`gram-bacteria`[i]),na.rm=TRUE)
        data_wide$FUNGI[i]<-mean(c(data_wide$AMF[i],data_wide$fungi[i]),na.rm=TRUE)
        data_wide$ALL_MICROBES[i]<-mean(c(data_wide$AMF[i],data_wide$bacteria[i],data_wide$fungi[i],data_wide$general[i],data_wide$`gram-bacteria`[i],data_wide$`gram+bacteria`[i],data_wide$protozoa[i]),na.rm=TRUE)
      }

  data_wide<-data_wide[,c(1,2,3,4,5,12,13,14)] # Trim down to broader groups
  
  data<-melt(data_wide,value.name="d13",id.vars=c("Sample_Name","Rhizosphere_Manipulation","Tree","Substrate_Type"))
  names(data)[5]<-"Microbial_Group"
  
  ## Add a column for tree ID
  data$Tree_ID<-substr(data$Sample_Name,1,5)
  
  ## Separate leaf and starch datasets
  leaf_PLFA<-filter(data,Substrate_Type=="L")
  starch_PLFA<-filter(data,Substrate_Type=="S")  
```

## **Leaf Substrate**: Step-wise AIC Model Selection for 13C-PLFA
```{r leaf_13C_PLFA_variables, include=FALSE}
  leaf_13C_PLFA<-leaf_PLFA$d13
  Exclusion<-leaf_PLFA$Rhizosphere_Manipulation
  Tree<-leaf_PLFA$Tree
  Microbe_Group<-leaf_PLFA$Microbial_Group
```

```{r leaf_13C_PLFA_stepAIC}
leaf_13CPLFA_AIC<-stepAIC(object=aov(lm(leaf_13C_PLFA~Exclusion*Tree*Microbe_Group)),direction="both",na.rm=TRUE)
```

```{r leaf_13C_PLFA_AIC_anova_table}
leaf_13CPLFA_AIC$anova
```

```{r leaf_13C_PLFA_AIC_extraction}
print(extractAIC(aov(lm(leaf_13C_PLFA~Exclusion+Tree+Exclusion*Tree))))
print(extractAIC(aov(lm(leaf_13C_PLFA~Exclusion+Tree+Exclusion*Tree+Microbe_Group))))
```

Model selection indicates that **rhizosphere exclusion, tree species, and exclusion x tree interaction** should be included in model explaining **13C-PLFA** in cores labeled with **leaf substrate**. Model that included microbial group was ~6 AIC points greater than the final model, so microbial group was not included in final analyses.

### Type II ANOVA Results
```{r leaf_13C_PLFA_ANOVA}
leafMod_13CPLFA<-lm(d13~Rhizosphere_Manipulation+Tree+Rhizosphere_Manipulation*Tree,data=leaf_PLFA)
Anova(leafMod_13CPLFA,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r leaf_nested_13C_PLFA}
leaf_PLFA_mixed_mod<-lme(d13~Rhizosphere_Manipulation+Tree+Rhizosphere_Manipulation*Tree,random=~1|Tree_ID,data=leaf_PLFA,na.action=na.omit,method="ML")
leaf_PLFA_fixed_mod<-lm(d13~Rhizosphere_Manipulation+Tree+Rhizosphere_Manipulation*Tree,data=leaf_PLFA,na.action=na.omit)
anova(leaf_PLFA_mixed_mod,leaf_PLFA_fixed_mod)
```

This output indicates that the AIC score for the mixed model is ~30 points lower, with a p-value of < 0.0001, a clear indicator that the mixed model is without a doubt the best model (tree ID explains a substantial amount of variation).

```{r leaf_PLFA_mixed_mod}
summary(leaf_PLFA_mixed_mod)
```

### Tukey's HSD Results
```{r leaf_13C_PLFA_TukeyHSD}
tx<-with(leaf_PLFA, interaction(Rhizosphere_Manipulation,Microbial_Group,Tree))
amod<-aov(d13 ~ tx, data=leaf_PLFA)
HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

## **Starch Substrate**: Step-wise AIC Model Selection for 13C-PLFA
```{r starch_13C_PLFA_variables, include=FALSE}
  starch_13C_PLFA<-starch_PLFA$d13
  Exclusion<-starch_PLFA$Rhizosphere_Manipulation
  Tree<-starch_PLFA$Tree
  Microbe_Group<-starch_PLFA$Microbial_Group
```

```{r starch_13C_PLFA_stepAIC}
starch_13CPLFA_AIC<-stepAIC(object=aov(lm(starch_13C_PLFA~Exclusion*Tree*Microbe_Group)),direction="both",na.rm=TRUE)
```

```{r starch_13C_PLFA_AIC_anova_table}
starch_13CPLFA_AIC$anova
```

```{r starch_13C_PLFA_AIC_extraction}
print(extractAIC(aov(lm(starch_13C_PLFA~Exclusion+Microbe_Group))))
print(extractAIC(aov(lm(starch_13C_PLFA~Exclusion+Microbe_Group+Tree))))
```

Model selection indicates that **rhizosphere exclusion, microbial group, and tree species** should be included in model explaining **13C-PLFA** in cores labeled with **starch substrate**. Model with only rhizosphere exclusion and microbial group and model with rhizosphere exclusion, microbial group, and tree species are separated by only 1 AIC point, so both models are included in "best family" of models. Will choose to include tree species as a parameter to match structure of CO2 Analyses.

### Type II ANOVA Results
```{r starch_13C_PLFA_ANOVA}
starchMod_13CPLFA<-lm(d13~Rhizosphere_Manipulation+Microbial_Group+Tree,data=starch_PLFA)
Anova(starchMod_13CPLFA,type="II") 
```

### Compare Nested Model to Fixed Effects Model
```{r starch_nested_13C_PLFA}
starch_PLFA_mixed_mod<-lme(d13~Rhizosphere_Manipulation+Microbial_Group+Tree,random=~1|Tree_ID,data=starch_PLFA,na.action=na.omit,method="ML")
starch_PLFA_fixed_mod<-lm(d13~Rhizosphere_Manipulation+Tree+Rhizosphere_Manipulation*Tree,data=starch_PLFA,na.action=na.omit)
anova(starch_PLFA_mixed_mod,starch_PLFA_fixed_mod)
```

This output indicates that the AIC score for the mixed model is ~30 points lower, with a p-value of < 0.0001, a clear indicator that the mixed model is without a doubt the best model (tree ID explains a substantial amount of variation).

```{r leaf_PLFA_mixed_mod}
summary(leaf_PLFA_mixed_mod)
```

### Tukey's HSD Results
```{r starch_13C_PLFA_TukeyHSD}
tx<-with(starch_PLFA, interaction(Rhizosphere_Manipulation,Microbial_Group,Tree))
amod<-aov(d13 ~ tx, data=starch_PLFA)
HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

# 13C Specific Respiration Models
```{r read_13C_specResp_data, include=FALSE}
  ## Read in data
  data<-read.csv("Final_Specific_Respiration_Data.csv")
  data$Mesh_Type<-as.factor(data$Mesh_Type)
  leaf<-filter(data,Label_Type=="L")
  starch<-filter(data,Label_Type=="S")
```

## **Leaf Substrate**: Step-wise AIC Model Selection for 13C Specific Respiration
```{r leaf_13C_specResp_variables, include=FALSE}
  leaf_13C_specresp<-leaf$Specific_Respiration_CO2_FAME
  Exclusion<-leaf$Mesh_Type
  Tree<-leaf$Tree_Species
```

```{r leaf_13C_specResp_stepAIC}
leaf_13C_specresp_AIC<-stepAIC(object=aov(lm(leaf_13C_specresp~Exclusion*Tree+Exclusion+Tree)),direction="both",na.rm=TRUE)
```

```{r leaf_13C_specResp_AIC_anova_table}
leaf_13C_specresp_AIC$anova
```

```{r leaf_13C_specResp_AIC_extraction}
print(extractAIC(aov(lm(leaf_13C_specresp~Exclusion*Tree))))
```

Model selection indicates that all parameters, i.e. **rhizosphere exclusion, tree species, and exclusion x tree interaction** should be included in model explaining **13C-specific respiration** in cores labeled with **leaf substrate**. 

### Type II ANOVA Results
```{r leaf_13C_specResp_ANOVA}
leaf_mod<-lm(Specific_Respiration_CO2_FAME~Mesh_Type*Tree_Species,data=leaf)
Anova(leaf_mod,type="II")
```

### Tukey's HSD Results
```{r leaf_13C_specResp_TukeyHSD}
tx<-with(leaf, interaction(Mesh_Type,Tree_Species))
amod<-aov(Specific_Respiration_CO2_FAME ~ tx, data=leaf)
HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

## **Starch Substrate**: Step-wise AIC Model Selection for 13C Specific Respiration
```{r starch_13C_specResp_variables, include=FALSE}
  starch_13C_specresp<-starch$Specific_Respiration_CO2_FAME
  Exclusion<-starch$Mesh_Type
  Tree<-starch$Tree_Species
```

```{r starch_13C_specResp_stepAIC}
starch_13C_specresp_AIC<-stepAIC(object=aov(lm(starch_13C_specresp~Exclusion+Tree+Exclusion*Tree)),direction="both",na.rm=TRUE)
```

```{r starch_13C_specResp_AIC_anova_table}
starch_13C_specresp_AIC$anova
```

```{r starch_13C_specResp_AIC_extraction}
print(extractAIC(aov(lm(starch_13C_specresp~Exclusion+Tree))))
print(extractAIC(aov(lm(starch_13C_specresp~Exclusion))))
print(extractAIC(aov(lm(starch_13C_specresp~Tree))))
print(extractAIC(aov(lm(starch_13C_specresp~1))))
```

Model selection indicates that the best model for **13C-specific respiration of starch substrate** includes **only the intercept**, but the model including **tree species** as a parameter is within the family of best models. 

### Type II ANOVA Results
```{r starch_13C_specResp_ANOVA}
starch_mod<-lm(Specific_Respiration_CO2_FAME~Mesh_Type+Tree_Species,data=starch)
Anova(starch_mod,type="II")
```

### Tukey's HSD Results
```{r starch_13C_specResp_TukeyHSD}
tx<-with(starch, interaction(Mesh_Type,Tree_Species))
amod<-aov(Specific_Respiration_CO2_FAME ~ tx, data=starch)
HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

# Background Respiration Models

## Step-wise AIC Model Selection for Background Respiration
```{r read_back_resp_data, include=FALSE}
data<-read.csv("Resp_Data_for_Spec_Resp.csv")
```

```{r back_resp_variables}
Resp<-data$Mean_Respiration
Tree<-data$Tree_species
Exclusion<-data$Exclusion
```

```{r back_resp_stepAIC}
back_resp_AIC<-stepAIC(object=aov(lm(Resp~Exclusion+Tree+Exclusion*Tree)),direction="both",na.rm=TRUE)
```

```{r back_resp_AIC_anova_table}
back_resp_AIC$anova
```

```{r back_resp_AIC_extraction}
print(extractAIC(aov(lm(Resp~Exclusion))))
print(extractAIC(aov(lm(Resp~Tree))))
print(extractAIC(aov(lm(Resp~Exclusion+Tree))))
```

Model selection indicates that the best model for **background respiration** includes **rhizosphere exclusion and tree species**. Model containing only rhizosphere exclusion is technically the best model, but the model containing both rhizosphere exclusion and tree species falls within the "family of best models".

### Type II ANOVA Results
```{r back_resp_ANOVA}
resp_mod<-lm(Mean_Respiration~Exclusion+Tree_species,data=data)
Anova(resp_mod,type="II")
```

### Tukey's HSD Results
```{r back_resp_TukeyHSD}
tx<-with(data, interaction(Exclusion,Tree_species))
amod<-aov(Mean_Respiration ~ tx, data=data)
HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

```{r back_resp_TukeyHSD_Exclusion}

amod<-aov(data$Mean_Respiration~as.factor(data$Exclusion))
TukeyHSD(amod)
```


# Background PLFA Models

## Step-wise AIC Model Selection for Background PLFA
```{r read_back_PLFA_data, include=FALSE}
data<-read.csv("Unlab_PLFA_for_Spec_Resp.csv")
```

```{r back_PLFA_variables}
FAME<-data$ln_FAME
Tree<-data$Tree
Exclusion<-data$Rhizosphere_Manipulation
Microbe<-data$Microbial_Group
```

```{r back_PLFA_stepAIC}
back_PLFA_AIC<-stepAIC(object=aov(lm(FAME~Exclusion*Tree*Microbe)),direction="both",na.rm=TRUE)
```

```{r back_PLFA_AIC_anova_table}
back_PLFA_AIC$anova
```

Model selection indicates that the best model for **background PLFA** includes **rhizosphere exclusion, tree species, and microbial group**, but no interaction terms.

### Type II ANOVA Results
```{r back_PLFA_ANOVA}
PLFA_mod<-lm(FAME~Exclusion+Tree+Microbe,data=data)
Anova(PLFA_mod,type="II")
```

### Tukey's HSD Results
```{r back_PLFA_TukeyHSD}
tx<-with(data, interaction(Exclusion,Tree,Microbe))
amod<-aov(FAME ~ tx, data=data)
HSD.test(amod, "tx", group=TRUE, console=TRUE)
```

# Background Specific Respiration Models

## Step-wise AIC Model Selection for Background Specific Respiration
```{r read_back_specresp_data, include=FALSE}
data<-read.csv("Unlab_Spec_Resp_CO2_and_PLFA_Data.csv")
```

```{r back_specresp_variables}
specResp<-data$Specific_Respiration_CO2_FAME
Exclusion<-data$Rhizosphere_Manipulation
Tree<-data$Tree
```

```{r back_specresp_stepAIC}
back_specresp_AIC<-stepAIC(object=aov(lm(specResp~Exclusion*Tree)),direction="both",na.rm=TRUE)
```

```{r back_specresp_AIC_anova_table}
back_specresp_AIC$anova
```

Model selection indicates that the best model for **background specific respiration** includes **rhizosphere exclusion and tree species** (no interaction term).

### Type II ANOVA Results
```{r back_specresp_ANOVA}
back_specresp_mod<-lm(specResp~Exclusion+Tree,data=data)
Anova(back_specresp_mod,type="II")
```

### Tukey's HSD Results
```{r back_specresp_TukeyHSD}
tx<-with(data, interaction(Exclusion,Tree))
amod<-aov(specResp ~ tx, data=data)
HSD.test(amod, "tx", group=TRUE, console=TRUE)
```
